Work Fragmentation in Developer Interaction Data

### 2016-08-10 paper_2.tex
pass

\begin{changedforreviewerlong}
While these studies already show the effect of work fragmentation can be hurtful, they are mostly observational studies over a limited time, and do not analyze in detail the use of tools from where evidence could be taken about work fragmentation. In the case of programmers the use of tools to create software is pervasive; software development environments (IDE) are the contexts where a significant proportion of the work is done and the interaction with them can mark moments of interruption \cite{BK16}.

Similarly, previous studies although explore the nature of activities of information workers, do not characterize them at a level which can be used to understand particular aspects of programming as a professional activity. According to Parnin and Rugaber \cite{PR12}, programmers must perform activities that challenge the memory, like keeping track of refactoring changes, navigate and debug unfamiliar code, learn new APIs, and understand new concepts; when these tasks are interrupted, it can be costly to recover the lost state. Some of these tasks require to invest in program comprehension activities, which can account from 54 to 95\% of the time spent on every working session \cite{MMLK14}.

@@

It was also required to fill the description for the events that indicate the activation or deactivation of the workbench of Eclipse, but as they were the only events with this error this task was trivial. Additionally, we needed to classify the events, which was a more complicated task, described below.

@@

@@

Certainly, work fragmentation is not a synonym of lower productivity. Multitasking can allow the worker to use dead periods of the project where there is a decrease of the required activities and do tasks related to other projects, increasing the efficiency and productivity \cite{ABV12}. At some level, switching between tasks can provide the needed arousal and alertness to improve the productivity, but higher levels of multitasking can be detrimental and cause a negative effect. Moreover, as the number of switches between tasks increases, the effectiveness degrades in a linear fashion \cite{AB12}. Also, it was observed that switching between projects might degrade the productivity levels of the programmer due to the context switches, so developers that work on many projects per day have lower productivity \cite{VBX16}.

Our results indicate a clear drop of productivity when facing interruptions: when we increase the number of interruptions, the productivity metrics tend to decrease. This agrees with the observational studies cited before, even if we can not relate our interruptions to actual multitasking.

>>>>

\begin{changedforreviewerlong}
While these studies already show the effect of work fragmentation can be hurtful, they are mostly observational studies over a limited time, and do not analyze in detail the use of tools from where additional evidence could be gathered to better understand work fragmentation. In the case of programmers the use of tools to create software is pervasive; integrated software development environments (IDEs) are tools where a significant proportion of the work is done and the interaction with them (or lack thereof) can mark moments of interruption \cite{BK16}.

Similarly, previous studies although they explore the nature of activities of information workers, do not characterize them at a level which can be used to understand particular aspects of programming as a professional activity. According to Parnin and Rugaber \cite{PR12}, programmers must perform activities that challenge the memory, such as keeping track of refactoring changes, navigate and debug unfamiliar code, learn new APIs, and understand new concepts; when these tasks are interrupted, it can be costly to recover the lost state. Some of these tasks require to invest time in program comprehension activities, which can account from 54 to 95\% of the time spent on every working session \cite{MMLK14}.

@@

We also had to fill the description for the events that indicate the activation or deactivation of the workbench of Eclipse, but as they were the only events with this error this task was trivial. Additionally, we needed to classify the events, which was a more complicated task, described below.

@@

\begin{changedforreviewerlong}
To ensure that a small subset of users do not dominate our dataset, we counted the number of sessions attributed to each UDC user. The third quartile was 5 sessions, the median 15, and the third quartile 38; the minimum was 1, and the maximum 270. These figures tell us that if some users have more sessions than others (which is to be expected), there is no major bias due to a small number of users concentrating the vast majority of the sessions.
\end{changedforreviewerlong}

@@

\begin{changedforreviewerlong}
Certainly, work fragmentation is not a synonym of lower productivity. Multitasking can allow the worker to use dead periods of the project where there is a decrease of the required activities and do tasks related to other projects, increasing the efficiency and productivity \cite{ABV12}. At some level, switching between tasks can provide the needed arousal and alertness to improve the productivity, but higher levels of multitasking can be detrimental and cause a negative effect. Moreover, as the number of switches between tasks increases, the effectiveness degrades in a linear fashion \cite{AB12}. However, a study by Monk found that when faced with a higher freqency of interruptions, subjects programming a VCR had a reduced resumption lag, indicating that they adapted to the interruption frequency \cite{monk2004effect}. Also, it was observed that switching between projects might degrade the productivity levels of the programmer due to the context switches, so developers that work on many projects per day have lower productivity \cite{VBX16}.

Our results indicate a clear drop of productivity when facing interruptions: when we increase the number of interruptions, the productivity metrics tend to decrease, but the most dramatic decrease happens for the first interruptions. This agrees with the observational studies cited before, particularly Monk's, even if we can not relate our interruptions to actual multitasking.
\end{changedforreviewerlong}
